Benchmarking FDR Methods

### 2017-09-14 simulation_studies/fdr_simulation_studies.Rmd
overhauled null simulations using /R/simulation-helpers.R; made du_ttest_sim() output generic for any test statistic, effect size, p value, and independent covariate

```{r}
m = 20000 # number of tests
M = 1000 # number of MC reps
```

@@

## Control for FDR at $\alpha$

Evaluate methods control for FDR at nominal $\alpha$ level ranging from [0.01, 0.1].

```{r}
# notes: convert this into R code
# - IHW: p-values from t-test, independent covariate
# - ASH: difference in means, independent covariate
# - BL: p-values from t-test, multiple covariates (with and without cov structure)
#   - lambda: 0.75, 0.85, 0.95
# - Scott: t-test statistic, independent covariate
#   - assumption of test statistic follow a standard normal. this is good here. 
# - BH: p-values only. 
# - locfdr: zz = t-test statistic, nulltype (try both). 
# - mixfdr: ? do we want to include?

```

>>>>

```{r}
m = 5000 # number of tests
M = 4 # number of MC reps

 # load helper functions
RPROJ <- list(PROJHOME = normalizePath(getwd()))
source(file.path(RPROJ$PROJHOME, "R/simulation-helpers.R"))
```

@@

## Control for FDR and FWER at $\alpha$

In this section, we will plot the number of discoveries 
as a function of $\alpha$ (should be 0). We will also evaluate
methods control for false discovery rate (FDR) and family-wise 
error rate (FWER) at nominal $\alpha$ level ranging from [0.01, 0.1].

```{r, load-libraries, message=FALSE}
library(dplyr)
library(ggplot2)
library(doParallel)
```

Additional packages that you should install are: 
```{r, eval=FALSE}
library(genefilter)

library(IHW)
library(ashr)
library(qvalue)
library(swfdr)
library(fdrtool)
  ## to use IHWpaper::scott_fdrreg wrapper, need to install FDRreg v2.0 (only available github):   
  ## devtools::install_github(repo= 'jgscott/FDRreg', subdir='R_pkg/') 
library(FDRreg)
```

Simulate an example "null" dataset. `H` is an indicator representing the true null (0) or alternative status (1). `test_statistic` is the t-test statistic, `effect_size` is the difference in means, 
`pval` is the $p$-value from the t-test, `ind_covariate` (in this case) is the pooled standard deviation ($\sigma$), which reduces to the unpooled sd in this case, 
and `SE` is the standard error ($\frac{\sigma}{\sqrt{n}}$)
```{r example-simulate-null-data}
n_samples = 20
n_groups = 2

sim_df <- du_ttest_sim(m=m, pi0=1, effect_size=0, 
                       n_samples=n_samples, n_groups = 2)
head(sim_df)
```

Run following on odyssey and save output as RDS: 
```{r evaluate-methods-null-data, eval=FALSE}
library(dplyr)
library(ggplot2)
library(doParallel)

workingPath <- "/net/irizarryfs01/srv/export/irizarryfs01_backed_up/share_root/shicks/projects/benchmark-fdr"
source(file.path(workingPath, "simulation-helpers.R"))

# set these to a smaller number temporarily 
m = 5000 # number of tests
M = 500 # number of MC reps

n_samples = 20
n_groups = 2

alphas = seq(0.01, 0.1, length.out = 10)

nCores <- 10
registerDoParallel(cores = nCores)
workers <- getDoParWorkers()
backend <- getDoParName()
version <- getDoParVersion()

df.out <- foreach(i = 1:M, .verbose=T) %dopar% {
  sim_df <- du_ttest_sim(m=m, pi0=1, effect_size=0, n_samples=n_samples, n_groups=2)
  sim_runner(sim = sim_df, alphas = alphas)
}

df.out <- dplyr::bind_rows(df.out)

saveRDS(df.out, file=file.path(workingPath, "nullSims-Results.RDS"))
```

Load and plot results from RDS file: 
```{r, eval=FALSE}
# workingPath <- "/net/irizarryfs01/srv/export/irizarryfs01_backed_up/share_root/shicks/projects/benchmark-fdr"
# df.out <- readRDS(file=file.path(workingPath, "nullSims-Results.RDS"))
df.out <- readRDS(file=file.path(RPROJ$PROJHOME, "R/nullSims-Results.RDS"))

# Number of discoveries as a function of $\alpha$
df.out %>% dplyr::group_by(method, alpha) %>% 
      dplyr::filter(!(method %in% c("ash", "scott", "locfdr"))) %>%
      dplyr::summarize(n_rejects = mean(n_rejects), FDR = mean(FDP), 
                       FWER = mean(FWER), FPR = mean(FPR)) %>% 
    ggplot(aes(x = alpha, y = n_rejects, color = method)) + 
         geom_line() + geom_abline(linetype="dashed") + 
         xlab(expression(bold(paste("Nominal ",alpha)))) + ylab("Number of Discoveries") +
         theme(axis.title = element_text(face="bold") )

# False discovery rate as a function of $\alpha$
df.out %>% dplyr::group_by(method, alpha) %>%
        # dplyr::filter(!(method %in% c("ash", "scott", "locfdr"))) %>%
      dplyr::summarize(n_rejects = mean(n_rejects), FDR = mean(FDP), 
                       FWER = mean(FWER), FPR = mean(FPR)) %>% 
      ggplot(aes(x = alpha, y = FDR, color = method)) + 
         geom_line() + geom_abline(linetype="dashed") + 
         xlab(expression(bold(paste("Nominal ",alpha)))) + ylab("FDR") +
         theme(axis.title = element_text(face="bold") )

# Family wise error rate as a function of $\alpha$
df.out %>% dplyr::group_by(method, alpha) %>% 
         #  dplyr::filter(!(method %in% c("ash", "scott"))) %>%
      dplyr::summarize(n_rejects = mean(n_rejects), FDR = mean(FDP), 
                       FWER = mean(FWER), FPR = mean(FPR)) %>% 
      ggplot(aes(x = alpha, y = FWER, color = method)) + 
         geom_line() + geom_abline(linetype="dashed") + 
         xlab(expression(bold(paste("Nominal ",alpha)))) + ylab("FWER") +
         theme(axis.title = element_text(face="bold") )
```
